Reliable error estimation for Sobol' indices

### 2016-03-31 replicated_sobol.tex
Changes to commit to replicated Sobol

\section{Problem Statement}
Denote by $\vx \in [0,1)^d$ a point with components $x_1,\dots,x_d$. Let $u$ be a subset of $\{1,\dots,d\}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, we define the following hybrid point: 
\[(\vx_u:{\vx'}_{-u}) := ({x'}_1,\dots,{x'}_{u-1},x_u,{x'}_{u+1},\dots,{x'}_d).\]
Let $f$ be the symbol representing the numerical model considered. We assume $f \in \mathcal{L}^2[0,1]^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
%These two quantities can be expressed as the following integrals:
%\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}, \]
%\[\sigma^2= \int_{[0,1)^{d}}f(\vx)^2 d\vx - \Big{(}\int_{[0,1)^{d}}f(\vx) d\vx\Big{)}^2 .\]
The uncertainty on $\vx$ is modeled by random variables such that $\vx \sim \mathcal{U}[0,1]^d$. Recall the Hoeffding decomposition of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \{1,\dots,d\}} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, by taking the variance of each side in Equation \ref{anova} we obtain the following decomposition of the variance:
\[ \sigma^2 = \sum \limits_{u \subseteq \{1,\dots,d\}} \sigma_v^2.\]
The problem of interest is the estimation of the quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad \text{ for } u = 1,\dots,d.\]
These quantities correspond to the partial sensitivity Sobol' indices often found written in the literature in their normalized form: $\underline{\tau}_u^2/\sigma^2$. When, $|u|=1$, $\underline{\tau}_u^2$ corresponds to the first-order Sobol' index that evaluate the main effect of the variable $\vx_u$. In \cite{}, Owen proposed to use the following expression for $\underline{\tau}_u^2$:
\begin{equation}\label{first_order_indice}
\underline{\tau}_u^2  =\int_{[0,1)^{2d-1}}(f(\vx) - \mu) (f(\vx_u:{\vx'}_{-u})-\mu) d\vx d{\vx'}_{-u},
\end{equation}
taking $\mu$ as introduced in Janon \textit{et al.}:
\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}. \]
The natural way to estimate these two quantities is for $(\vx_i,\vx'_i) \stackrel{iid}{\sim} [0,1)^{2d}$ via:
\begin{align}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=1}^n (f(\vx_i) - \widehat{\mu}) (f(\vx_{i,u}:{\vx'}_{i,-u})-\widehat{\mu}),\label{first_order_indice_estimator}\\
\nonumber
\widehat{\mu} & = \frac{1}{2n} \sum \limits_{i=1}^n f(\vx_i) +f(\vx_{i,u}:{\vx'}_{i,-u}).
\end{align}
\bigskip

>>>>

\section{Problem Statement}
Denote by $\vx \in [0,1)^d$ a point with components $x_1,\dots,x_d$. Let $u$ be a subset of $\{1,\dots,d\}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, we define the following hybrid point: 
\[(\vx_u:{\vx'}_{-u}) := ({x'}_1,\dots,{x'}_{u-1},x_u,{x'}_{u+1},\dots,{x'}_d).\]
Let $f$ be the symbol representing the numerical model considered. We assume $f \in \mathcal{L}^2[0,1]^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
%These two quantities can be expressed as the following integrals:
%\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}, \]
%\[\sigma^2= \int_{[0,1)^{d}}f(\vx)^2 d\vx - \Big{(}\int_{[0,1)^{d}}f(\vx) d\vx\Big{)}^2 .\]
The uncertainty on $\vx$ is modeled by random variables such that $\vx \sim \mathcal{U}[0,1]^d$. Recall the Hoeffding decomposition of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \{1,\dots,d\}} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{d-|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, by taking the variance of each side in Equation \ref{anova} we obtain the following decomposition of the variance:
\[ \sigma^2 = \sum \limits_{u \subseteq \{1,\dots,d\}} \sigma_v^2.\]
The problem of interest is the estimation of the quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad \text{ for } u = 1,\dots,d.\]
These quantities correspond to the partial sensitivity Sobol' indices often found written in the literature in their normalized form: $\underline{\tau}_u^2/\sigma^2$. When, $|u|=1$, $\underline{\tau}_u^2$ corresponds to the first-order Sobol' index that evaluate the main effect of the variable $\vx_u$. In \cite{}, Owen proposed to use the following expression for $\underline{\tau}_u^2$:
\begin{equation}\label{first_order_indice}
\underline{\tau}_u^2  =\int_{[0,1)^{2d-1}}(f(\vx) - \mu) (f(\vx_u:{\vx'}_{-u})-\mu) d\vx d{\vx'}_{-u},
\end{equation}
taking $\mu$ as introduced in Janon \textit{et al.}:
\[\mu= \frac{1}{2} \int_{[0,1)^{2d-1}}f(\vx)+f(\vx_u:{\vx'}_{-u}) d\vx d{\vx'}_{-u}. \]
The natural way to estimate these two quantities is for $(\vx_i,\vx'_i) \stackrel{iid}{\sim} [0,1)^{2d}$ via:
\begin{align}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=1}^n (f(\vx_i) - \widehat{\mu}) (f(\vx_{i,u}:{\vx'}_{i,-u})-\widehat{\mu}),\label{first_order_indice_estimator}\\
\nonumber
\widehat{\mu} & = \frac{1}{2n} \sum \limits_{i=1}^n f(\vx_i) +f(\vx_{i,u}:{\vx'}_{i,-u}).
\end{align}
\bigskip
